---
title: "Time Series Assignment"
subtitle: 'DSTI A20 Cohort'
author: "Ariel Nataf"
output:
  pdf_document: default
---

Importing libraries
```{r, results='hide', message=FALSE}
#install.packages("readxl")
#install.packages("writexl")
library(readxl)
library(writexl)
library(forecast)
library(ggplot2)
library(vars)
```

Loading the data 
```{r}
data = read_excel("/Users/arielnataf/Desktop/DSTI/TimeSeries/Elec-train.xlsx")
```


# Forcasting without using outdoor temperature

frequency: 1 day = 96 quarters
start: 01/01/2010 01:15 -> 6th quarter of the day
end: 2/16/2010 23:45 -> 96th quarter of the 47th day

```{r, fig.show='hide'}
ts<-ts(data$`Power (kW)` ,start=c(1,6),end=c(47,96),freq=96)

ts_train = window(ts, start=c(1,6),end=c(46,96))
ts_test = window(ts, start=c(47,1),end=c(47,96))
plot(ts)
```

The time series appears to be cyclical over a 96 period day period

## Holt-Winter Forcasting

the serie has a trend and a seasonal pattern, HW should be a good model.


```{r}
#For hs(), the max frequency is 24. We have to aggregate the values.
ts24 = aggregate(ts, nfrequency = 24)/4

# We don't have an increase of amplitude
# so we are only looking at addititive seasonality
fit1=hw(ts24,seasonal='additive',h=24)
autoplot(tail(ts24,72)) +
autolayer(fit1,series='HW add.',PI=FALSE)
```
The forecasting looks good, but **this solution doesn't give us 96 rows.**

## Seasonal ARIMA Forcasting

First looking at an auto ARIMA
```{r, fig.show='hide'}
fit_auto=auto.arima(ts_train,lambda = "auto")

prev_arima=forecast(fit_auto,h=96)

#autoplot(prev_arima)

```

> With auto.arima:
>
> ARIMA(5,0,0)(0,1,0)[96] 
>
>sigma^2 estimated as 0.4246:  log likelihood=-4272.53
>
> AIC=8557.07   AICc=8557.09   BIC=8595.29

With lag 96, the ACF and PACF are simply not readable.
I won't be attempting setting manually ARIMA

## Neural Network Auto-Regression (NNAR) Forecasting

```{r, results='hide', message=FALSE}
fit_nn=nnetar(ts_train,t=96)
prevNN=forecast(fit_nn,h=96)
```

## Selecting forecast

We are not considering the HW solution because it couldn't perform the 96 predictied values per day expected.

```{r}
plot(ts_test,xlim=c(47,48))
lines(ts_test,lty=2)
lines(prev_arima$mean,col=2)
lines(prevNN$mean,col=3)

legend('topleft',col=1:3,lty=1,legend=c('test','SARIMA','NN'))
```

We are looking at the RMSE

```{r}
rmse <- function(serie_test,prev) {return=sqrt(mean((serie_test-prev$mean)^2))}

sprintf('RMSE: • AUTO ARIMA: %f • NN: %f ',
        rmse(ts_test, prev_arima), rmse(ts_test, prevNN))
```

NN has better results than ARIMA! We will use this solution.

# Forcasting using outdoor temperature

We now have an additional covariate Temp (C°).

```{r}
colnames(data) <- c("Timestamp","Power","Temp")
#simpler to refer to objects with simple characters

ts2<-ts(data[c("Power","Temp")],start=c(1,6),end=c(47,96),freq=96)

ts2_train = window(ts2, start=c(1,6),end=c(44,96))
ts2_test = window(ts2, start=c(45,1),end=c(47,96))
```

## Time series regression forecasting

```{r, results='hide', message=FALSE}
?tslm
fit=tslm(Power~Temp+trend+season,data=ts2)
summary(fit)
CV(fit)
```
All the features seem relevant

```{r, fig.show='hide'}
checkresiduals(fit,test=FALSE,plot=TRUE)
```

```{r, fig.show='hide'}
plot(pacf(fit$residuals))

```

```{r}
tmp=fit$residuals
#fit2=Arima(tmp,order=c(13,0,0))
#checkresiduals(fit2)
#plot(pacf(fit2$residuals))

```

## VAR forecasting

```{r, results='hide', message=FALSE}
VARselect(ts2_train, lag.max=150, type="const")
```
Thanks to AIC, we choose order 103

```{r}
var <- VAR(ts2_train, p=103,type = "const")
prev_var=forecast(var,h=96,PI=FALSE)
plot(prev_var)
```


## Selecting forecast


```{r}
plot(ts2_test[,'Power'] ,xlim=c(45,48))
lines(ts2_test[,'Power'],lty=2)
lines(prev_var$forecast$Power$mean ,col=2)

legend('topleft',col=1:3,lty=1,legend=c('test','VAR'))
```


```{r}
rmse <- function(serie_test,prev) {return=sqrt(mean((serie_test-prev$mean)^2))}

sprintf('RMSE: • VAR: %f ', rmse(ts2_test[,"Power"], prev_var$forecast$Power))

```

# Generating and Saving Final Forecast


```{r, results='hide', message=FALSE}

fit_nn=nnetar(ts,t=96)
prevNN_results=forecast(fit_nn,h=96)
```

```{r}
var_results <- VAR(ts2, p=103,type = "const")
prev_var_results=forecast(var_results,h=96,PI=FALSE)

df <- data.frame(univariate = prevNN_results,
                 covariate = prev_var_results$forecast$Power)
df = subset(df, select = -c(3:8))
colnames(df) <- c("univariate", "covariate")
print(head(df))
```
```{r}
write_xlsx(df, "./Ariel_Nataf.xlsx")
```

